---
title: "Social Survey"
author: "Elisabet Mendoza"
date: "11/13/2017"
output:
  html_document: default
---

```{r include=FALSE,echo=FALSE}
require(tigerstats)
require(tidyverse)
source("Import_GSS_2014_haven.R")
```
<!-- We recommend you use git branches -- create a separate branch for each major section i.e. Introduction, Methods, Results, Inferential Results, and Discussion and Conclusions - leave the YAML block untouched until the final merge where you put in your own name !-->

# Introduction
Since 1972, every two years a social survey is conducted to the general public to see how society changes. We will be looking at the data from the 2014 general social survey to find if if there is a relationships between gun ownership and other variables. What is the possible relationships between gun ownership and race, sex, political views, opinions on capitol punishment and belief in god?
<!--In this section you explain what you are trying to show.  Where did the data come from?  What is the research or other question you are trying to answer?!-->   

####This study deals with questions about possible relationships between gun ownership and race, sex, political views, opinions on capitol punishment and belief in god. The data is from the 2014 general social survey.



# Methods
I will look at the data if the following charts. I will be comparing the relationships between gun ownership and race, sex, political views, opinions on capitol punishment and belief in god. The data will be compared and I will see both numerical and graphical results. Finally I will do a chi-square test to look at the inferential results.

<!--Decide on your methods:  use "variable analysis" or other appropriate descriptors.  Make sure to choose at least one graphical method and at least one numerical method.!-->

#Results

<!--Divide this section into two sub-sections:  One for your descriptive  results and one for your inferential results.!-->

## Descriptive Results

### Graphical Descriptive Results

<!--Graphical results here.  Make sure to show your code.  Provide appropriate labels for axes, giving units if possible, and provide a good title for the graph, too.  Use the graphical results to describe the patterns if any that exist in the data as focused toward the research question!-->

```{r}
source("makeTables.R")
```
```{r}
barchartGC(god_table,type="percent", main = "Belief in god and Gun ownership")
```

All the groups have a consistent pattern. All of the groups had a vote of about 30% "yes" and 70% "no" on gun ownership

```{r}
barchartGC(race_table,type="percent", main = "Race and Gun ownership")
```

In all groups there is a larger amount of people that vote "no" on gun ownership. White vote "yes" more often than black or other, but there are still more votes for "no" across the chart.

```{r}
barchartGC(sex_table,type="percent", main = "Sex and Gun ownership")
```

In this chart we can see that in both groups the vote for "no" is still dominant. We can also see that more males votes for "yes" than females did.

```{r}
barchartGC(polviews_table,type="percent", main = "Political views and Gun ownership")
```

Here we can see that all groups are very different. We see a kind of pattern the more conservative, the more vote "yes" to gun ownership, the more liberal the less vote "yes".

```{r}
barchartGC(cappun_table,type="percent", main = "Capitol punishment and Gun ownership")
```

In this chart we can see that he groups are very different. Although there is more votes on "no" for gun ownership, there is a higher amount of "yes" votes in the favor group than in the oppose group.

In all charts we can see that votes for "no" over powers the "yes" vote in gun ownership.

### Numerical Descriptive Results

<!--Numerical results go here. Use the numerical results to describe the patterns if any that exist in the data as focused toward the research question!-->
```{r}
rowPerc(god_table)
```

Looking at the numerical results between Belief in god and gun ownership we can see that all groups answered the same regardless of there believe in god.

```{r}
rowPerc(race_table)
```

Looking at the numerical results between race and gun ownership we can see that 39% of white said "yes" to gun ownership while only 17% of black and 14% of other said "yes".Even though the majority of all groups still say "no"

```{r}
rowPerc(sex_table)
```

Looking at the numerical results between sex and gun ownership we can see that 40% of males said "yes" to gun ownership while only 27% of females said "yes". Even though the majority answered "no" there are still more male that responded "yes" in the survey.

```{r}
rowPerc(polviews_table)
```

Looking at the numerical results between political view and gun ownership we can see that even though the majority in all groups answered "no" to gun ownership, we can see that the conservative group has the highest "yes" answers with 42%, followed by the moderate group with a 33%, last the liberal with only 25% that answered "yes".

```{r}
rowPerc(cappun_table)
```

Looking at the numerical results between favor or oppose in capital punishment and gun ownership we can see that the group in favor with 41% answer "yes". While only 21% of those who oppose answered "yes".

# Inferential Results

<!--State hypothesis clearly.  Make sure your discussion of the inferential test covers all the aspects that the test output produces, such as test statistic, p-value etc.  Make a decision about the null hypothesis, explain the assumptions on which the selected test/procedure was based, and why the chosen procedure satisfys the assumptions and is appropriate to answer the research question!-->
I will do a chi-square test to look at the inferential results. The chi-square test will tell us with exactitude if all of this variables have a relationship between each of them individually and gun ownership.

#####Belief in God and gun ownership
```{r}
chisqtestGC(god_table)
```
H0: Gun ownership is independent of belief in God.
Ha: Gun ownership and belief in God are not independent.

The P value of 0.97 is very close to the degrees of freedom from the chi-square test which is 2. Therefore, based on the data, we will have to agree with the null hypothesis. The null hypothesis states that there is not a direct relationship between ones belief in God and gun-ownership.

#####Race and gun ownership
```{r}
chisqtestGC(race_table)
```
H0: Gun ownership is independent of race.
Ha: Gun ownership and race are not independent.

Based of the p-value being low at approximately 0, it appears that we have to reject the null hypothesis. The null hypothesis therefore is rejected, and there is a direct relationship between race and gun-ownership.

#####Sex andgun ownership
```{r}
chisqtestGC(sex_table)
```
H0: Gun ownership is independent of sex.
Ha: Sex and Gun ownership are related (not independent).

The P-value is low at approximately 0, therefore, we have to reject the null hypothesis which states there is no direct relationship between gun ownership and sex. However, in this case, we can see there is a direct relationship as females tend to be more opposed to gun ownership then men.

#####Political view and gun ownership
```{r}
chisqtestGC(polviews_table)
```
H0: Gun ownership is independent of political views.
Ha: Gun ownership and political views are related (not independent).

The p-value is at approximately 0 which means that we have to reject the null hypothesis. The null hypothesis predicted that there would be more of the moderate political group that would be in favor of gun ownership. However, that is not the case as they highest YES rate came from the conservative group. Therefore, we would agree that there is a direct relationship between ones political views and gun ownership.

#####Capital punishment and gun ownership
```{r}
chisqtestGC(cappun_table)
```
H0: Gun ownership is independent of opinion on capital punishment.
Ha: Opinion on capital punishment and gun ownership are related (not independent).

Based on the p-value of 0, it appears that we would have to reject the null hypothesis. The null hypothesis states that there is no direct relationship between ones views on capital punishment and gun ownership. Therefore, we see there is a direct relationship between ones views on capital punishment and gun ownership.
# Discussion and Conclusion

<!--Discussion and conclusion here.  If you found a relationship be sure to consider whether the relationship occurs because one of the variables causes the other, or whether they perhasps are related for some other reason.  Watch the chapter 6 videos from the GeorgeTown videos collection.!-->
After viewing the results graphically, numerically, and running chi-square tests on the data, we can conclude the following:

Race and gun ownership are related. Sex and gun ownership are related. Political views and gun ownership are related. Belief in capital punishment and gun ownership are related. Belief in God and gun ownership are related.

If we were to make a profile of those that voted YES on gun ownership they might be WHITE, MALE, CONSERVATIVE, in FAVOR of capital punishment, and MAYBE believe or ACTUALLY believe in GOD.

If we were to make a profile of those that voted NO on gun ownership they might be BLACK or OTHER, FEMALE, LIBERAL or MODERATE, OPPOSE capital punishment, and DO NOT believe in GOD.

So without taking any other variables into account, there is a relationship between race, sex, political beliefs, beliefs on capital punishment, belief in God and gun ownership. Many participants in these groups voted NO on gun ownership but it seems as though they are influenced by their specific group. It would be interesting to research a generation gap on gun ownership. Would there be a relationship with age and gun ownership? The GSS data has over 3000 different variables which means there are many different factors to compare.